---
title: "Fitting Exercise"
---

First to install packages
```{r}
library(here)
library(dplyr)
library(ggplot2)
library(tidymodels)
```

Then to load data
```{r}
data_location <- here::here("fitting-exercise", "Mavoglurant_A2121_nmpk.csv")
df <- read.csv(data_location)
```

Plotting outcome (DV) as a function of time colored by dose and using ID as a grouping factor
```{r}
p1 <- ggplot(df, aes(x=TIME, y=DV, group=ID, color=DOSE)) + geom_line()
plot(p1)
```

remove all OCC=2 entries, keep only OCC=1
```{r}
df2 <- subset(df, (OCC == 1))
```

Exclude all Time=0, found sum of DV for each ID, create a new data frame with only time 0 and a new variable Y that contains amount of drug delivered over the whole experiment
```{r}
df3 <- subset(df2, (TIME != 0))
df4 <- df3 %>% group_by(ID) %>% summarise(Y = sum(DV))
df5 <- subset(df2, (TIME == 0))
df6 <- left_join(x=df5, y=df4, by=c("ID"))
```

Keep only variables Y, Dose, Age, Sex, Race, WT, HT. Convert Race and Sex to factors
```{r}
df7 <- select(df6,Y, DOSE, AGE, SEX, RACE, WT, HT)
df7$SEX <- as.factor(df7$SEX)
df7$RACE <- as.factor(df7$RACE)
```

Plots depicting:
dose and Y colored by Age
Y and weight colored by Sex
boxplot of Y by sex
boxplot of Y by race
```{r}
summary(df7)

p2 <- ggplot(df7, aes(x=DOSE, y=Y, color=AGE)) + geom_point()
plot(p2)

p3 <- ggplot(df7, aes(x=WT, y=Y, color=SEX)) + geom_point()
plot(p3)

p4 <- ggplot(df7, aes(x=SEX, y=Y)) + geom_boxplot()
plot(p4)

p5 <- ggplot(df7, aes(x=RACE, y=Y)) + geom_boxplot()
plot(p5)

hisDOSE <- ggplot(df7, aes(x=DOSE)) + geom_histogram()
plot(hisDOSE)

hisAGE <- ggplot(df7, aes(x=AGE)) + geom_histogram()
plot(hisAGE)

hisWT <- ggplot(df7, aes(x=WT)) + geom_histogram()
plot(hisWT)

hisHT <- ggplot(df7, aes(x=HT)) + geom_histogram()
plot(hisHT)
```

fit a linear model to Y using Dose and calculate RMSE/R2
```{r}
lm_mod1 <- linear_reg()
lm_fit1 <- lm_mod1 %>% fit(Y ~ DOSE, data = df7)
lm_fit1
pred1 <- predict(lm_fit1, df7) %>% bind_cols(df7)
met1 <- metrics(pred1, truth = Y, estimate = .pred)
met1
```
In this model dose has a coefficient of 58 which seems like a high number to me. (whenever my collaborators run models we get numbers like 1.2, although those are probably somewhat different models.) The R2 is .516 which is also decent in my opinion. I still have some issues figuring out how to interpret the RMSE and MAE though.


fit a linear model to Y using all predictors and calculate RMSE/R2
```{r}
lm_mod2 <- linear_reg()
lm_fit2 <- lm_mod2 %>% fit(Y ~ DOSE * AGE * SEX * WT * HT * RACE, data = df7)
lm_fit2
pred2 <- predict(lm_fit2, df7) %>% bind_cols(df7)
met2 <- metrics(pred2, truth = Y, estimate = .pred)
met2
```
This one has a lower RMSE/MAE and a higher R2, so I can conclude that it is probably a better model than the first.


fit a linear model to SEX using Dose, and one using all predictors, and calculate ROC-AUC
```{r}
lm_mod3 <- logistic_reg()
lm_fit3 <- lm_mod3 %>% fit(SEX ~ DOSE, data = df7)
lm_fit3
pred3 <- predict(lm_fit3, df7, type = "prob") %>% bind_cols(df7)
roc_auc(pred3, truth = SEX, .pred_1)

lm_mod4 <- logistic_reg()
lm_fit4 <- lm_mod4 %>% fit(SEX ~ DOSE * Y * AGE * WT * HT * RACE, data = df7)
lm_fit4
pred4 <- predict(lm_fit4, df7, type = "prob") %>% bind_cols(df7)
roc_auc(pred4, truth = SEX, .pred_1)
```

The roc_auc is 1 for the last model which is super interesting. I guess that means that all the data put together are very good at predicting sex. This makes sense given that a lot of the things in the model are correlated with sex (height/weight, and doeses tend to be different for men/women). In model 3, dose alone had an ROC of .6, so it's probably a decent predictor of sex.

random seed:
```{r}
rngseed = 1234
```

remove race as a variable
```{r}
df8 <- df7 %>% select(-RACE)
```

set seed and split data 75:25
```{r}
set.seed(rngseed)

data_split  <- initial_split(df8, prop = 3/4)

train <- training(data_split)
test <- testing(data_split)
```

two linear models for Y (continuous outcome of interest)
Model 1: dose as only pred
Model 2: all predictors
```{r}
set.seed(rngseed)
lm_mod5 <- linear_reg()
lm_fit5 <- lm_mod5 %>% fit(Y ~ DOSE, data = train)
lm_fit5
pred5 <- predict(lm_fit5, train) %>% bind_cols(train)
met5 <- metrics(pred5, truth = Y, estimate = .pred)
met5

lm_mod6 <- linear_reg()
lm_fit6 <- lm_mod6 %>% fit(Y ~ DOSE * AGE * SEX * WT * HT, data = train)
lm_fit6
pred6 <- predict(lm_fit6, train) %>% bind_cols(train)
met6 <- metrics(pred6, truth = Y, estimate = .pred)
met6

NM <- null_model(mode = "regression", engine = "parsnip")
NM_fit <- NM %>% fit(Y ~., data = train)
NM_fit
pred7 <- predict(NM_fit, train) %>% bind_cols(train)
met7 <- rmse(pred7, truth = Y, estimate = .pred)
met7
```

I set the random seed correctly, however my values are different from the ones we're supposed to get. I got 948 for the null which is correct, 703 for model 1 (correct value 702), and 578 for model 2 (correct value 627).
This does mean that model 2 is currently performing best as it has the lowest rmse

Now I need to do a 10-fold CV
```{r}
folds <- vfold_cv(train, v = 10)
folds

lm_wf1 <- workflow() %>% add_model(lm_mod5) %>% add_formula(Y ~ DOSE)
lm_fit_cv1 <- lm_wf1 %>% fit_resamples(folds)
lm_fit_cv1
collect_metrics(lm_fit_cv1)

lm_wf2 <- workflow() %>% add_model(lm_mod5) %>% add_formula(Y ~ DOSE * AGE * SEX * WT * HT)
lm_fit_cv2 <- lm_wf2 %>% fit_resamples(folds)
lm_fit_cv2
collect_metrics(lm_fit_cv2)

NM_wf <- workflow() %>% add_model(NM) %>% add_formula(Y ~ .)
NM_fit_cv <- NM_wf %>% fit_resamples(folds)
NM_fit_cv
collect_metrics(NM_fit_cv)
```

Well the rmse for the null model was similar but different at 943, SE at 66.2. RMSE for the second model with the full formula was much worse at 2003, while rmse for the Y~DOSE model was 687.

Now I need a different random seed
```{r}
set.seed(347)
folds2 <- vfold_cv(train, v = 10)
folds2

lm_wf3 <- workflow() %>% add_model(lm_mod5) %>% add_formula(Y ~ DOSE)
lm_fit_cv3 <- lm_wf3 %>% fit_resamples(folds2)
lm_fit_cv3
collect_metrics(lm_fit_cv3)

lm_wf4 <- workflow() %>% add_model(lm_mod5) %>% add_formula(Y ~ DOSE * AGE * SEX * WT * HT)
lm_fit_cv4 <- lm_wf4 %>% fit_resamples(folds2)
lm_fit_cv4
collect_metrics(lm_fit_cv2)

NM_wf2 <- workflow() %>% add_model(NM) %>% add_formula(Y ~ .)
NM_fit_cv2 <- NM_wf2 %>% fit_resamples(folds2)
NM_fit_cv2
collect_metrics(NM_fit_cv2)
```

Null model is now at 938, similar but slightly different
second model is still at 2003
First model is now at 684.
Similar, but slightly different to the other random seed.