---
title: "Fitting Exercise"
---

First to install packages
```{r}
library(here)
library(dplyr)
library(ggplot2)
```

Then to load data
```{r}
data_location <- here::here("fitting-exercise", "Mavoglurant_A2121_nmpk.csv")
df <- read.csv(data_location)
```

Plotting outcome (DV) as a function of time colored by dose and using ID as a grouping factor
```{r}
p1 <- ggplot(df, aes(x=TIME, y=DV, group=ID, color=DOSE)) + geom_line()
plot(p1)
```

remove all OCC=2 entries, keep only OCC=1
```{r}
df2 <- subset(df, (OCC == 1))
```

Exclude all Time=0, found sum of DV for each ID, create a new data frame with only time 0 and a new variable Y that contains amount of drug delivered over the whole experiment
```{r}
df3 <- subset(df2, (TIME != 0))
df4 <- df3 %>% group_by(ID) %>% summarise(Y = sum(DV))
df5 <- subset(df2, (TIME == 0))
df6 <- left_join(x=df5, y=df4, by=c("ID"))
```

Keep only variables Y, Dose, Age, Sex, Race, WT, HT. Convert Race and Sex to factors
```{r}
df7 <- select(df6,Y, DOSE, AGE, SEX, RACE, WT, HT)
df7$SEX <- as.factor(df7$SEX)
df7$RACE <- as.factor(df7$RACE)
summary(df7)
```

