---
title: "Future Exercise"
---

```{r}
library("dslabs")
str(gapminder)
summary(gapminder)
data(gapminder)

africadata <- subset(gapminder, (continent == "Africa"))
#creates a new dataframe from a subset of the old one where the continent is Africa
str(africadata)
#Checks to make sure I have the correct number of observations

inf_LE <- data.frame(africadata$infant_mortality, africadata$life_expectancy)
#creates a new dataframe from these two columns of the africadata dataframe
#this one is for infant mortality and Life expectancy
str(inf_LE)
#This checks to make sure I got it right

pop_LE <- data.frame(africadata$population, africadata$life_expectancy)
#This creates a new dataframe for population and life expectancy data
str(pop_LE)
#This checks that I have everything right

plot(inf_LE)
#plots this dataframe on a scatter plot
#plot(pop_LE$africadata.population, pop_LE$africadata.life_expectancy, log = "x")
#above is one option for generating the other scatter plot
plot(pop_LE, log = "x")
#here is a neater version
#the log function is where you can specify which axis should be in log, in this case population
```
My hypothesis for the streaks of data we see is that they are individual countries over time
```{r}
year_inf <- data.frame(africadata$year, africadata$infant_mortality)
#this creates a matrix with year and infant mortality
year_inf[is.na(year_inf)] <- "A"
#this line rewrites the dataset, changing all NA's to A
missing <- subset(year_inf, (africadata.infant_mortality == "A"))
#this makes a new dataset with the subset function that pulls out all the years with A as a value
#This didn't work with NA which is why I changed it to A. Not the most elegant solution but it works
str(missing)
#this line is for checking

year2000 <- subset(africadata, (year == "2000"))
#this creates a new data.frame with all the data from the year 2000 from africadata
str(year2000)
#this is for checking

plot(year2000$infant_mortality, year2000$life_expectancy)

plot(year2000$population, year2000$life_expectancy, log = "x")
```
These are the updated plots for the year 2000!

```{r}
fit1 <- lm(year2000$life_expectancy ~ year2000$infant_mortality)
summary(fit1)
fit2 <- lm(year2000$life_expectancy ~ year2000$population)
summary(fit2)
```
I used the help command to figure out what I needed to input into the lm command. It said the format was response ~ predictor. It also looked like I could set several responses and predictors.

This section is contributed by KATHERINELORUSSO
```{r}
library(dplyr)
library(ggplot2)
str(us_contagious_diseases)
#str function was used to see how many observations and variables (16065 observations & 6 variables)
summary(us_contagious_diseases)
#summary was used to view the variables and statistical values of the variables. 

GAdata <- us_contagious_diseases %>% filter(state == "Georgia")
#I wanted to only look at Georgia data, so I used the filter function and named it GAdata.
GAmumps <- GAdata %>%filter(disease=="Mumps") %>%select(-weeks_reporting, -population)
#I then wanted to only look at Mumps, so I used the filter function. I also used the select(-) because I just wanted to have the count and year values. 
summary(GAmumps)

ggplot(GAmumps, aes(year,count))+
    geom_point()+
    scale_x_continuous(breaks=seq(1968,2002, by=3))
#I used the ggplot function to set year on the x axis and count on the y axis and the geom_point() function to create a scatter plot. It is seen that there was an upwards trend in cases from until around 1998, a major spike in 1989, and a decrease in the following years. 
model <- glm(count~year, poisson, data=GAmumps)
summary(model)
#I used a generalized linear model to see if there is an association between year and count. After doing research, the poisson distribution was the best fit because this model is counts over time. There is a significant negative correlation of -0.011243 between year and count of Mumps in Georgia. 
``` 
