---
title: "Complex Data analysis"
---

I ran a Jaccard analysis on a set of Data of mine looking at _Salmonella_ populations at 3 creek sites. I went sampling after it had been dry for 3 days, and also after rain. I want to know if the populations differ signifigantly between the wet/dry samplings. A Jaccard analysis measures beta diversity and compares every sample to every other sample producing essentially, a triangular matrix. I have been procrastinating on figureing out how to analyze and present this data for several months, so we're doing it now. 
Based on some readings I have done, the typical test for something like this is a PERMANOVA.
My current plan: load my matrix, load a vector of dry samples and a vector of wet samples, then figure out how to use these vectors and the matrix for the PERMANOVA and then plot whatever the output is. I hope this counts as complex data.

Load Libraries
```{r}
library(ggplot2) #for making plots
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(here) #to set paths
library(vegan) #this is a package that allows me to do a PERMANOVA I think
```
This loads all the libraries I need for this project

Load Data
```{r}
data_location <- here::here("data-exercise", "CC_CSS_Jaccard2.csv")
jmatrix <- read.csv(data_location, row.names = 1)
#note to future self, do not populate row.names with the () version, use row.names = for it to work
```
This cell loads the output of my Jaccard analysis
Note: I still need to redo the Jaccard with the last of my CSS files, so this is not final data.


Load Vectors
```{r}
dry.wet_location <- here::here("data-exercise", "Samples_wet_dry.csv")
dry.wet <- read.csv(dry.wet_location, header = TRUE)

```
This cell loads a file that denotes which samples are wet or dry

```{r}
result <- adonis2(jmatrix ~ Wet.Dry, dry.wet, permutations = 999, method = "jaccard")
#this should run the permanova test using the adonis2 formula in the vegan package
print(result)
```
I did something! I have no idea what the printed result fully means though